\boldmath Supersymmetric Dirac-Born-Infeld Axionic Inflation and Non-Gaussianity

### 2018-12-02 SUSY_DBI_referee_reply.tex

2.2) A similar analysis should be done for the combination  $a_+$. After eq. (3.26), the paper states "$a_+$ undergoes fast roll.... thus we suppress $a_+$". This point should be clarified. My understanding is that this sentence means that $a_+$ has a much larger mass than $a_-$, so we can consistently set $a_+$ = 0. This should be shown explicitly, by computing explicitly the mass of this field (in terms of the parameters in the model) and showing that it is indeed greater than H.\\

\noindent {\it Reply\\
This is one of the central points of the analysis, i.e., decomposition of the potential into slow roll vs fast roll where the slow roll potential
 depends  only on $a_-$ and the remaining part of the potential is controlled by $a_+$ and is fast roll. The reason we did not elaborate
 too much on  it in this paper is because this topic was  dealt with at great length in our previous paper  ``Evidence for Inflation in an Axion Landscape,''  JHEP {\bf 1803}, 121 (2018). In this work it was shown that $a_+$ has a much larger mass than $a_-$ and further
 that the $a_+$ rolls an order of magnitude or more  faster than $a_-$ as shown by blue lines (fast roll) vs red lines (slow roll) in Fig.1
 and Fig.2 of JHEP {\bf 1803}, 121 (2018).
However, we are happy to elaborate on this point in this work. The mass of $a_+$ is controlled largely by $W_s$ and $W_{sb}$ makes
 significantly smaller contribution to $a_+$. In the analysis we choose $\mu=M_P$ where $M_P$ is the Planck mass while
  the mass scales that enter $W_{sb}$ are significantly smaller than $M_P$.
Thus to a good approximation the mass of $a_+$ is controlled by $\mu$ alone and a simple analysis using $W_s$ shows that
$m_{a_+}\simeq 2 \sqrt 2 \mu$. We have checked  that $m_{a_+}\simeq 2 \sqrt 2 \mu$ is larger than $H$
for the entire parameter space we investigate.
We note that $\mu$ does not enter in the $a_-$ mass as can be seen in Eqs.~(3.31)-(3.34) and
choose parameter space of $W_{sb}$ such that $m_{a_-}$ is smaller that $m_{a_+}$. We exhibit these results graphically in Fig.~(5) in the
revised paper. Here Fig.~(5a) is a histogram of $H$ values which shows that $m_{a_+}$ is always larger than $H$.
Fig.~(5b) is a histogram of $m_{a_-}$ values which shows that $m_{a_-}$ is always significantly smaller than $m_{a_+}$ for the parameter space chosen.
In addition to Fig.~(5)
we have added a further discussion of this in text in the revised version of the paper
to explain this point further. This is done at the end of the paragraph after Eq.~(3.27).
}

@@

3) There are different shapes of non-gaussianity (the equilateral and the local shapes being the most studied ones), and the limits on non-gaussianity strongly depend on the shape. By shape, we mean the dependence of the bispectrum on the ratios $k_2/k_1$ and
$k_3/k_1$. The paper cites [64,65,66]. Eq. 11 of [64], which is also eq. 1 of [65], and eq. 17 of [66] defines the so called local shape. This is different from the shape of non-Gaussianity obtained in DBI models, which is very close to the equilateral shape (see for instance Section 2.1 of the Planck paper 1502.01592). Moreover the works [64,65,66], although very important, are rather outdated by now. So, this part of the paper should be updated, and the value of $f_{NL}$ should be compared to the current limits (see 1502.01592) or future forecasts on equilateral non-Gaussianity, not the local one.

\noindent {\it Reply\\ 
We agree with the referee
that our references to the analysis of non-Gaussianity from the Planck experiment needed updating. We have
done so in the revised version.  The changes made appear after Eq.(5.12) and near the end of section 6. We have included
the reference to 1502.01592 and specified the current limits on $f_{NL}^{equil}$.
}

4) I don't understand eq.(6.1). The conventional way to find the end
of inflation is to find the moment at which $\ddot a = 0$ (second derivative of the
scale factor with respect to time). The number of e-folds during inflation is
typically defined from $a(N ) \equiv a_{end} e^{-N}$ , which gives  $\frac{dN}{dt} = -H$ (where $a_{end}$ 
is the value of the scale factor at the end of inflation, and H is the Hubble
rate). In this way, $N = 0$ at the end of inflation, while N is positive at
early times during inflation. How is eq. (6.1) used to determine the end of
inflation ?  What is the purpose of setting $\theta_n = \frac{1}{16}$?  Why don't the authors 
find from their numerical evolutions the point at which $\ddot a = 0$ ?\\
\noindent {\it Reply\\ 
We agree with the referee  regarding the conventional way to find  the end of inflation.
For the Dirac-Born-Infeld the numerical integration is time intensive especially for the very large scan
involving $0.5\times 10^6$ parameter sets that we carried out.  For this reason we approximated the end of inflation the way we did
which is much faster to implement than the $\ddot {a}=0$ constraint. However, we have taken time to 
carry out the analysis using the $\ddot{a}=0$ constraints and our analysis shows that the results of our approximate
procedure and the conventional ones agree for 97\% of the parameter points. In any case we have replaced the old 
Figs (1) and (2) with new ones using $\ddot {a}=0$ constraint. Further, we have replaced Eq.(6.1) with the new constraint.
}

5) The scan shown in Figure 1 contains no real information, unless one
specifies what is varied among the different points. The only information I
could find on this regard is "we allow $\alpha_1$ to vary" at the end of page 12.
How is it varied ? (in which range ? Is it a uniform sampling in linear units
? In log units ?) What is chosen for the other parameters of the model
? ($\mu$,$\lambda$,A,T). How is the correct normalization of the scalar perturbations
enforced? (this is typically done by fixing the overall scale of the scalar
potential). Are subleading terms in $\frac{1}{T}$  important ? The scan strategy should
be completely described, so that any reader, if interested, could precisely reproduce those figures.\\

\noindent {\it Reply\\
We agree that the distribution used to generate the dataset from Figs.(1, 2) is important to be
mentioned in the paper. We added a complete specification to the caption of Fig.(1).
Since $a_-$ is a pNGB, $\mu$ and $\lambda$ do not enter in the slow-roll Lagrangian, as can be seen in Eqs.~(3.30) to (3.35).
}

@@

6) The paper motivates the model from the trans-Planckian problem of axion inflation (the fact that we must take the axion scale 
$f > M_p$ in the original potential of natural inflation). It would be very interesting to provide also a scan of the values of f obtained for the various realizations presented in figures 1 and 2. Does this scan extend in the sub-Planckian region ? (the scan should be easily done, once the above questions 2.4 and 5 are answered).\\

\noindent {\it Reply\\
We agree with the suggestion of the referee. We have added a new figure 4 and a short discussion of it at the end of section 6 where we have presented a scatter plot of $f / M_\text{P}$ vs $n_s$ and $f / M_\text{P}$ vs $r$ for the data sets of figures 1 and 2. Here we see that $f$
 lies in the sub-Planckian region for a signifiant set of parameter points.}\\

{\it In the above we have replied to all of the comments and suggestion of the referee. We have revised the manuscript to
take account of them. We are submitting the revised manuscript  for publication in JHEP.}

### 2018-12-02 SUSY_DBI.tex
Fix typos and style (#19)
# Changes

 * Fixes typos, style, and updates fig5 with units.

   because $|W_{sb}| << |W_s|$. In the analysis of this paper there is just one pair of axion field so we have only $a_+$ and $a_-$ 

@@

  \begin{figure}
    \centering
    \includegraphics[width=1.0\textwidth]{figs/fig1.pdf}
    \caption{Left panel: A plot of the ratio $r$ of the tensor to scalar power spectrum vs the
      scalar spectral index $n_s$ for the two field DBI model.
      The blue region enclosed by the blue line is the one allowed by experiment in the 2$\sigma$
      range. $89\%$ of the simulated points are consistent with Planck constraints as shown by green
      points while the red points lie outside the 68\% CL contour.
          Right panel: A plot of non-Gaussianity parameter $f^{equil}_{NL}$ as
      a function of $\alpha_1$ for the same data set as in the left plot. As in the left panel
      the green dots are parameter points which satisfy the Planck constraints
      on $r$ and $n_s$ and lie in the blue region and the red dots are parameter points which are
      outside the experimentally allowed region on $r$ and $n_s$.
          Points are distributed according to the following:
      $\alpha_1 \sim X_{0, \infty}$, $\alpha_2 = \alpha_3 = 0$,
      $T = 10^{-12} M_\text{P}^4$,
      $f \sim X_{0, \infty} M_\text{P}$,
      $\tilde\beta \sim X_{0, \infty}$,
      $m = 6$, % G_7 = G_8 = ... = 0
      ${\cal{G}}_1 = {\cal{G}}_2 = {\cal{G}}_3 = 0$, ${\cal{G}}_4 = 1$,
      ${\cal{G}}_5 \sim {\cal{G}}_6 \sim \mathcal{U}\left\{-1, 1\right\} \times X_{0, \infty}$,
      $N_{\rm pivot} \sim \mathcal{U}\left(50, 60\right)$,
      $a_{-, 0} \sim \mathcal{U}\left(0, 2\pi\right) \times f$,
      $\dot a_{-, 0} \sim \mathcal{U}\left(-1, 1\right)
          \times \left(2 \sqrt{T}/\left(\sqrt{2} \sqrt{\alpha_1} + 2\right)^{1/2}\right)$
      where $X_{0, \infty} = \mathcal{U}\left\{X_{0, 1}, 1 / X_{0, 1}\right\}$,
      $X_{0, 1} = \mathcal{U}\left(0, 1\right)$, $\mathcal{U}$ refers to a uniform distribution,
      and the distribution of $\dot a_{-, 0}$ is chosen such that the expression under the
      square root in Eq.~(\ref{eq.axionLagrangian.I}) is positive.
      Points are further filtered such that the mass of the inflaton $m_{a_-} < 0.1 M_\text{P}$.}
    \label{fig1}
  \end{figure}

@@

  \begin{figure}
  	\centering
  	\includegraphics[width=1.0\textwidth]{figs/fig3.pdf}
  	\caption{The effect of $\alpha_2$ (left panel) and $\alpha_3$ (right panel) on non-Gaussianity for a parameter point 
	taken from Fig.~(\ref{fig1}) which is consistent with 
	Planck data and has the  largest non-Gaussianity among the set of parameter points 
  in Fig.~(\ref{fig1}). Specifically, here $f = 0.429 M_\text{P}$, $a_{-, 0} = 0.233 M_\text{P}$, $\dot a_{-, 0} = -0.118 M_\text{P}^2$, $m = 6$, ${\cal{G}}_1 = {\cal{G}}_2 = {\cal{G}}_3 = 0$, ${\cal{G}}_4 = 1$, ${\cal{G}}_5 = -1.53$, ${\cal{G}}_6 = 563$, $\alpha_1 = 1207$, $\tilde \beta = 0.802$, $T = 10^{-12} M_\text{P}^4$. One
	finds that
	 the effect is not sufficient to obtain a significantly larger value of $f_\text{NL}$.}
  	\label{fig3}
  \end{figure}

@@

  \begin{equation}
    \ddot{a} = a H^2(1 + \frac{\dot H}{H^2}) = 0.
  \end{equation}
  An analytic expression similar to Eq.~(\ref{eq.axionLagrangian.II}), however,  cannot be derived  for  non-zero $\alpha_2$ and $\alpha_3$. In this case,  equations for the auxiliary fields $F_k$ are  solved numerically,  and the Lagrangian in terms of $\phi$ (Eq.~(\ref{Lag.23})) is
  also evaluated numerically for each set of values for ($a_-$, $\dot a_-$).  
   This process is significantly slower, therefore, in this case the simulations are only done for $29$ sets of parameter values as can be seen in Fig.~(\ref{fig3}). 
    The time of horizon exit is then found by counting $N_\text{pivot}$ e-foldings back from the end of integration, where $N_\text{pivot}$ is varied between $50$ and $60$.

@@

  \begin{figure}
    \centering
    \includegraphics[width=1.0\textwidth]{figs/fig4.pdf}
    \caption{
      Left panel: A plot of $f / M_{\rm P}$ as a function of $n_s$ which is in the allowed range of
      experiment and $N_{\rm pivot}$ in the range $50-60$ for the same data set as in
      Fig.~(\ref{fig1}).
      Right panel: The same as the left panel except that the plot is of $f / M_{\rm P}$ as a
      function of $r$.
      We note that the plots show that there are a large number of parameter points where
      $f / M_{\rm P} < 1$ consistent with all the experimental data. Also note there are
      experimentally consistent points with $f/M_{\rm P} \in \left[0.1, 1\right]$.}
    \label{fig4}
  \end{figure}

  \begin{figure}
    \centering
    \includegraphics[width=1.0\textwidth]{figs/fig5.pdf}
    \caption{
      Left panel: A histogram of the values of Hubble parameter at horizon exit for the same
      data set in Fig.~(\ref{fig1}) for points that are consistent (green) and not consistent (red)
      with experimental constraints on $r$ and $n_s$.
      Right panel: The same as the left panel except the histogram of the inflaton $a_-$ mass is
      shown instead.}
    \label{fig5}
  \end{figure}

>>>>

   because $|W_{sb}| << |W_s|$. In the analysis of this paper there is just one pair of axion fields so we have only $a_+$ and $a_-$

@@

  \begin{figure}
    \centering
    \includegraphics[width=1.0\textwidth]{figs/fig1.pdf}
    \caption{Left panel: A plot of the ratio $r$ of the tensor to scalar power spectrum vs the
      scalar spectral index $n_s$ for the two field DBI model.
      The blue region enclosed by the blue line is the one allowed by experiment in the 2$\sigma$
      range. $85\%$ of the simulated points are consistent with Planck constraints as shown by green
      points while the red points lie outside the 68\% CL contour.
          Right panel: A plot of non-Gaussianity parameter $f^{equil}_{NL}$ as
      a function of $\alpha_1$ for the same data set as in the left plot. As in the left panel
      the green dots are parameter points which satisfy the Planck constraints
      on $r$ and $n_s$ and lie in the blue region and the red dots are parameter points which are
      outside the experimentally allowed region on $r$ and $n_s$.
          Points are distributed according to the following:
      $\alpha_1 \sim X_{0, \infty}$, $\alpha_2 = \alpha_3 = 0$,
      $T = 10^{-12} M_\text{P}^4$,
      $f \sim X_{0, \infty} M_\text{P}$,
      $\tilde\beta \sim X_{0, \infty}$,
      $m = 6$, % G_7 = G_8 = ... = 0
      ${\cal{G}}_1 = {\cal{G}}_2 = {\cal{G}}_3 = 0$, ${\cal{G}}_4 = 1$,
      ${\cal{G}}_5 \sim {\cal{G}}_6 \sim \mathcal{U}\left\{-1, 1\right\} \times X_{0, \infty}$,
      $N_{\rm pivot} \sim \mathcal{U}\left(50, 60\right)$,
      $a_{-, 0} \sim \mathcal{U}\left(0, 2\pi\right) \times f$,
      $\dot a_{-, 0} \sim \mathcal{U}\left(-1, 1\right)
          \times \left(2 \sqrt{T}/\left(\sqrt{2} \sqrt{\alpha_1} + 2\right)^{1/2}\right)$
      where $X_{0, \infty} = \mathcal{U}\left\{X_{0, 1}, 1 / X_{0, 1}\right\}$,
      $X_{0, 1} = \mathcal{U}\left(0, 1\right)$, $\mathcal{U}$ refers to a uniform distribution,
      and the distribution of $\dot a_{-, 0}$ is chosen such that the expression under the
      square root in Eq.~(\ref{eq.axionLagrangian.I}) is positive.
      Points are further filtered such that the mass of the inflaton $m_{a_-} < 0.1 M_\text{P}$.}
    \label{fig1}
  \end{figure}

@@

  \begin{figure}
  	\centering
  	\includegraphics[width=1.0\textwidth]{figs/fig3.pdf}
  	\caption{The effect of $\alpha_2$ (left panel) and $\alpha_3$ (right panel) on non-Gaussianity for a parameter point 
	taken from Fig.~(\ref{fig1}) which is consistent with 
	Planck data and has the  largest non-Gaussianity among the set of parameter points 
in Fig.~(\ref{fig1}). Specifically, here $f = 0.429 M_\text{P}$, $a_{-, 0} = 0.233 M_\text{P}$, $\dot a_{-, 0} = -0.118 \sqrt{T}$, $m = 6$, ${\cal{G}}_1 = {\cal{G}}_2 = {\cal{G}}_3 = 0$, ${\cal{G}}_4 = 1$, ${\cal{G}}_5 = -1.53$, ${\cal{G}}_6 = 563$, $\alpha_1 = 1207$, $\tilde \beta = 0.802$, $T = 10^{-12} M_\text{P}^4$. One
	finds that
	 the effect is not sufficient to obtain a significantly larger value of $f_\text{NL}$.}
  	\label{fig3}
  \end{figure}

@@

  \begin{equation}
    \ddot{a} = a H^2 \left(1 + \frac{\dot H}{H^2} \right) = 0.
  \end{equation}
  An analytic expression similar to Eq.~(\ref{eq.axionLagrangian.II}), however,  cannot be derived  for  non-zero $\alpha_2$ and $\alpha_3$. In this case,  equations for the auxiliary fields $F_k$ are  solved numerically,  and the Lagrangian in terms of $\phi$ (Eq.~(\ref{Lag.23})) is
  also evaluated numerically for each set of values for ($a_-$, $\dot a_-$).  
   This process is significantly slower, therefore, in this case the simulations are only done for $29$ sets of parameter values as can be seen in Fig.~(\ref{fig3}). 
    The time of horizon exit is then found by counting $N_\text{pivot}$ e-foldings back from the end of integration, where $N_\text{pivot}$ is varied between $50$ and $60$.

@@

  \begin{figure}
    \centering
    \includegraphics[width=1.0\textwidth]{figs/fig4.pdf}
    \caption{
      Left panel: A plot of $f / M_{\rm P}$ as a function of $n_s$ which is in the allowed range of
      experiment and $N_{\rm pivot}$ in the range $50-60$ for the same data set as in
      Fig.~(\ref{fig1}).
      Right panel: The same as the left panel except that the plot is of $f / M_{\rm P}$ as a
      function of $r$.
      We note that there are
      experimentally consistent points with $f/M_{\rm P} \in \left[0.1, 1\right]$.}
    \label{fig4}
  \end{figure}

  \begin{figure}
    \centering
    \includegraphics[width=1.0\textwidth]{figs/fig5.pdf}
    \caption{
      Left panel: A histogram of the values of Hubble parameter at horizon exit for the same
      data set as in Fig.~(\ref{fig1}) for points that are consistent (green) and not consistent (red)
      with experimental constraints on $r$ and $n_s$.
      Right panel: The same as the left panel except the histogram of the inflaton $a_-$ mass is
      shown instead.}
    \label{fig5}
  \end{figure}
